\large{Vipassana for Hackers} \\ \Huge{The Proposal} \\ \large\textit{Version 0.1}

### 2024-06-27 the-proposal/the-proposal.tex
up to design & methods

There has been limited, unprofessional writing on the topic of the dangers of meditation.
\cite{ingram2020mastering,https://doi.org/10.1080/03060497.1983.11084578}
Even though these writings focus on Christian ideals and meditative concepts
(such as ``The Dark Night of The Soul''),
which have no relationship to Vipassana meditation whatsoever,
they do arise in discussions of Vipassana meditation from time to time.
As a result, it is a specific aim of this study to determine if there are
psychological or physical dangers presented by prolonged meditation practice.

@@

\subsection{Preliminary Research Review}
- summary of related research
- strengths and weaknesses
- justification
--- what hasn't been done by others?
--- why is this research necessary?

@@

1. ``complete'', standardized, global, multilingual, 100\% free
2. mundane (sleep) vs. supramundane (total eradication of suffering)

3. who is steven deobald

@@

@@

>>>>

There has been limited, unprofessional writing on the topic of the dangers of meditation.
\cite{ingram2020mastering,https://doi.org/10.1080/03060497.1983.11084578}
Even though these writings focus on Christian ideals, struggles of faith,
and meditative concepts (such as ``The Dark Night of The Soul''),
which have no relationship to meditation whatsoever,
they do arise in discussions of meditation from time to time.
As a result, it is a specific aim of this study to determine if there are
psychological or physical dangers presented by prolonged meditation practice.

@@

\subsection{Preliminary Research Review}

Related research, as mentioned previously, falls on the spectrum of exploratory
research conducted on ``expert'' (50,000 hours of practice or more) meditators
in various traditions or acute, quantitative research conducted on novice
(0 - 10,000 hours) or intermediate (10,000 - 50,000 hours) meditators.

The strength of each form of study is demonstrated in its results.
Exploratory research has given us a clear understanding that the brains of
very experienced meditators function in a manner very different from the
societal mean.
Quantitative research with a narrow hypothesis has provided repeated
statistical evidence that meditation practices do change the minds and bodies
of the participant, and that those changes are measurable.
Statistical research has also taught us that the effects of many
meditation practices become more pronounced with a prolonged practice period.

Longitudinal studies have largely not been conducted in the various fields
of meditation research.
Where they have been conducted, ``longitudinal'' may only represent a period
of nine months.
Generally, these longitudinal studies focus on larger groups of novice meditators.
While useful in its own right, this is equivalent to studying hobby pianists
for long periods of time: none of the subjects will ever master the piano.

True longitudinal study of an individual (or individuals) will allow for a
new form of exploratory research, which will itself uncover new questions
we might ask of meditation.
We can then probe with further statistical research once we know what
questions we want to ask of the technique or techniques under study.

@@

Vipassana is a ``complete'' system of meditation, based on two premises:

1. that every mental and emotional formation in the consciousness of an individual
corresponds to an equivalent physical formation in the bodily sensations
of that individual and

2. that the totality of physical, bodily sensations are open to exploration by
the method of self-directed human awareness

These premises can be experientially validated, and are thus known to be true.
As with all experience-oriented factual information found in the fields of
consciousness studies, we must operate based on the assumption that if a
third party wishes to validate these premises, they can and will do so
experientially.
This presents a difficulty, but not an insurmountable one, as we'll see below.

\subsubsection{Standardization and Access}

Vipassana meditation is a globally standardized meditation practice.
It is taught based on the same instructions across the globe,
and has been translated into dozens of languages, covering the vast
majority of global society.
Those instructions are taught verbally, via audio and video recording,
and do not require the participant to be literate in the language they speak.
Introductory courses in Vipassana meditation are all identical, covering
a 10-day period in which meditators participate in complete silence,
devoid of external input and confounding variables.

These introductory courses (as with all Vipassana meditation courses)
are taught completely free of charge, making access universal.

The only limitation to access is the mental and physical fitness of
individuals applying to take Vipassana meditation courses.
Course management will, at their discretion, request someone not to
take a course if they suffer from mental illness, for example.

\subsubsection{Purpose of Practice}

The intended purpose of Vipassana meditation is inherently supramundane:
``the total eradication of suffering.''
While this purpose may appear unreasonable or unattainable to skeptics,
it is not scientific to judge the potential outcomes of such a meditation
based on our assumptions.
We must instead test the methodology of Vipassana meditation to see
how it holds up to these claims, and to what degree.

Conversely, the purpose of many contemporary meditation techniques
(particularly those studied in existing literature) is inherently mundane:
to improve sleep, to reduce anxiety, to reduce PTSD, and so on.
While some of these effects may be visible in the study of Vipassana
meditation, they are not the end goal of the technique under study.

@@

Instruction in Vipassana meditation has been available outside of
Burma / Myanmar since 1969.
However, it has taken five decades of proliferation at the hands of
volunteers to make Vipassana meditation accessible to large portions of
the globe.
Even though there are permanent Vipassana Centres (teaching schools) in
many countries worldwide, there is still extremely limited presence in
Africa and the Middle East, though there is widespread interest there.

10-Day introductory courses in Vipassana meditation have long waitlists
in all of these Centres.
Demand for instruction in Vipassana meditation far outstrips supply.

Vipassana instruction is clearly on a path of growth, at present.
With a globally-accessible technique of meditation which is increasingly
discussed within the scope of popular culture, it is an appropriate
time to study its effects for the benefit of the global population.
\cite{marriage2024theretreat,muraskin2024perils,sacchet2024advanced}

@@

\subsubsection{Why Steven Deobald?}
